\documentclass[12pt]{article}

% Packages
\usepackage{geometry}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{float}
\usepackage{listings}
\usepackage{xcolor}

% Code listing style
\lstdefinestyle{codeStyle}{
    basicstyle=\ttfamily\small,
    breaklines=true,
    frame=single,
    numbers=left,
    numberstyle=\tiny,
    tabsize=4,
    showstringspaces=false
}

\geometry{margin=1in}

\title{
Hands-On MIS Projects\\
Chapter 1 - Information Systems in Global Business Today
}

\author{
Parsa Mohebian
}

\date{\today}


\begin{document}

\maketitle

\tableofcontents

\newpage


\section{Introduction}

This report analyzes management information system problems and demonstrates
how information systems improve decision making using transactional sales data.


\section{Management Decision Problem 1-7: Snyder's of Hanover}

\subsection{Problem Summary}

Snyder's financial department depended heavily on spreadsheets and manual processes
for collecting and reporting financial information. Every month, the financial analyst
collected spreadsheets from more than 50 departments worldwide, consolidated the data,
and created the company's profit-and-loss statement manually. This process was slow,
difficult to maintain, and vulnerable to errors.

\subsection{Impact on Business Performance}

The spreadsheet-based system negatively affected business performance.

\begin{itemize}
    \item The financial analyst spent significant time collecting and consolidating
    spreadsheets instead of performing financial analysis.
    
    \item Manual data entry increased the possibility of errors and inconsistent reports.
    
    \item Financial information was delayed because reports required manual preparation.
    
    \item The company faced higher operational costs because employees performed
    repetitive administrative tasks.
    
    \item The system was difficult to scale as the company expanded globally.
\end{itemize}


\subsection{Impact on Management Decision Making}

The existing system limited management decision making because managers did not have
fast access to accurate and updated financial information.

Delayed reporting made it harder to identify financial problems, compare department
performance, and respond quickly to changes in business conditions. Incorrect or
outdated information could result in poor decisions regarding budgeting, production,
and resource allocation.


\subsection{Recommended Information System Solution}

Snyder's should implement an integrated financial information system based on a
centralized database. The system should allow departments to enter financial data
directly into a shared platform and automatically generate standardized reports.

The system should include automated reporting, security controls, and management
dashboards. An Enterprise Resource Planning (ERP) system could integrate financial
data with other business functions such as sales, inventory, and production.

This solution would reduce manual work, improve reporting accuracy, provide faster
access to information, and support better management decisions.


\section{Management Decision Problem 1-8: Dollar General}

\subsection{Problem Summary}

Dollar General operates discount stores where minimizing costs is essential.
However, the company does not have an automated inventory tracking system at
each store. Managers cannot accurately verify incoming deliveries, which has
resulted in inventory errors and increasing merchandise losses.

The company needs to determine whether an information system investment can
improve inventory accuracy, reduce losses, and support better operational
decisions.

\subsection{Decisions Before Investing in an Information System}

Before investing in an inventory management system, Dollar General should
evaluate several important factors.

\begin{itemize}

\item \textbf{Identify business goals:}
The company should define the problems it wants to solve, such as reducing
inventory losses, improving inventory accuracy, and ensuring product availability.

\item \textbf{Determine required system features:}
The company should decide whether it needs barcode scanning, real-time inventory
tracking, automatic stock updates, inventory alerts, and reporting tools.

\item \textbf{Evaluate costs and benefits:}
Management should compare implementation costs, including hardware, software,
and training, against expected benefits such as lower losses and improved
operational efficiency.

\item \textbf{Plan implementation:}
The company must decide how the system will be introduced, how employees will be
trained, and how existing inventory data will be transferred.

\item \textbf{Consider security requirements:}
The company should establish rules for protecting inventory data and controlling
access to the system.

\end{itemize}

Dollar General should implement an automated inventory management system using
barcode scanning and a centralized database. This system would provide accurate
inventory information, reduce merchandise losses, improve replenishment decisions,
and increase overall operational efficiency.


\section{Database Analysis Project 1-9}

\subsection{Dataset Description}

The project uses a transactional marketplace database stored in SQLite format.
The database contains raw sales transaction data that can be used to analyze
product demand, regional sales performance, seller performance, and monthly sales
trends.

The database contains five main tables:

\begin{itemize}
    \item \texttt{orders}: contains order identification numbers, order dates,
    and customer contact information.
    
    \item \texttt{order\_items}: contains the products sold in each order,
    seller identifiers, product identifiers, and item prices.
    
    \item \texttt{products}: contains product identifiers, product categories,
    and product weight information.
    
    \item \texttt{sellers}: contains seller identifiers and seller state
    information.
    
    \item \texttt{order\_reviews}: contains customer review scores and review
    comments.
\end{itemize}

The main analysis used the \texttt{orders}, \texttt{order\_items},
\texttt{products}, and \texttt{sellers} tables. The \texttt{order\_reviews}
table was not required for the three main sales trend questions, but it could be
used in future analysis to study customer satisfaction.

The dataset included 96,478 orders, 110,197 order item records, 32,216 products,
2,970 sellers, and 39,462 reviews. The order dates ranged from September 15,
2016 to August 29, 2018.

\begin{table}[H]
\centering
\begin{tabular}{lr}
\toprule
Table & Number of Records \\
\midrule
\texttt{orders} & 96,478 \\
\texttt{order\_items} & 110,197 \\
\texttt{products} & 32,216 \\
\texttt{sellers} & 2,970 \\
\texttt{order\_reviews} & 39,462 \\
\bottomrule
\end{tabular}
\caption{Database Tables and Record Counts}
\end{table}


\subsection{Data Preparation}

The original dataset was provided as a SQLite database file. Because the project
requires database querying and reporting, SQLite was used to inspect the database
structure and run SQL queries. The database was also exported to CSV files so the
tables could be opened in spreadsheet or database software if needed.

A Python script was created to export each SQLite table into a separate CSV file.
The exported files were:

\begin{itemize}
    \item \texttt{orders.csv}
    \item \texttt{order\_items.csv}
    \item \texttt{products.csv}
    \item \texttt{sellers.csv}
    \item \texttt{order\_reviews.csv}
\end{itemize}

The main sales analysis was performed by joining the order, product, and seller
tables into a single SQL view named \texttt{sales\_analysis\_base}. This view
created a structure similar to the textbook sales database.

\begin{table}[H]
\centering
\begin{tabular}{ll}
\toprule
Textbook Field & Dataset Field Used \\
\midrule
Store identification number & \texttt{seller\_id} \\
Sales region & \texttt{seller\_state} \\
Item number & \texttt{product\_id} \\
Item description & \texttt{product\_category\_name} \\
Unit price & \texttt{price} \\
Units sold & Count of records in \texttt{order\_items} \\
Sales period & Month extracted from \texttt{timestamp} \\
\bottomrule
\end{tabular}
\caption{Mapping Textbook Sales Fields to Dataset Fields}
\end{table}

A missing data check was also performed on the joined sales view. The analysis
found no missing product categories, seller states, or prices in the final
analysis table.


\subsection{Sales Analysis Queries}

SQL queries were used to answer the three management questions in the assignment.
The queries joined the transactional tables and summarized the data by product
category, seller region, seller, and month.

The first query group identified which products should be restocked by ranking
product categories and individual products by units sold and total revenue. Since
the dataset did not include inventory quantity, high sales volume was used as a
proxy for high customer demand.

The second query group identified stores and regions that may benefit from
promotional campaigns. In this dataset, \texttt{seller\_id} represented the
store or seller, and \texttt{seller\_state} represented the sales region. Regions
with low revenue and low sales volume were treated as candidates for additional
marketing support.

The third query group analyzed monthly sales trends. Since the dataset did not
include a direct discount field, monthly revenue and units sold were used to
identify stronger and weaker sales periods. Stronger months were considered
better candidates for full-price selling, while weaker months were considered
better candidates for discounts or promotional campaigns.


\subsection{Restocking Analysis}

The first management question asks which products should be restocked. The dataset
does not include current inventory levels or stock quantities. Therefore, units sold
were used as a proxy for customer demand. Product categories with the highest sales
volume were treated as the strongest candidates for restocking.

The analysis showed that the highest-demand product categories were
\texttt{cama\_mesa\_banho}, \texttt{beleza\_saude}, \texttt{esporte\_lazer},
\texttt{moveis\_decoracao}, and \texttt{informatica\_acessorios}.

\begin{table}[H]
\centering
\begin{tabular}{lrr}
\toprule
Product Category & Units Sold & Total Revenue \\
\midrule
\texttt{cama\_mesa\_banho} & 10,953 & 1,023,434.76 \\
\texttt{beleza\_saude} & 9,465 & 1,233,131.72 \\
\texttt{esporte\_lazer} & 8,431 & 954,852.55 \\
\texttt{moveis\_decoracao} & 8,160 & 711,927.69 \\
\texttt{informatica\_acessorios} & 7,644 & 888,724.61 \\
\bottomrule
\end{tabular}
\caption{Top Product Categories by Units Sold}
\end{table}

Based on this analysis, managers should prioritize restocking these high-demand
categories. The category \texttt{beleza\_saude} is especially important because it
has both high sales volume and the highest total revenue among the top categories.
This suggests that keeping this category available could have a significant impact
on sales performance.


\subsection{Promotional Campaign Analysis}

The second management question asks which stores and sales regions would benefit
from a promotional campaign and additional marketing. In this dataset,
\texttt{seller\_id} was used as the store identifier, and \texttt{seller\_state}
was used as the sales region.

The analysis compared sales performance by region using units sold, number of
sellers, total revenue, and average price. Regions with low total revenue were
identified as possible candidates for promotional campaigns or additional
marketing support.

\begin{table}[H]
\centering
\begin{tabular}{lrrr}
\toprule
Region & Units Sold & Number of Sellers & Total Revenue \\
\midrule
AM & 3 & 1 & 1,177.00 \\
PA & 8 & 1 & 1,238.00 \\
SE & 10 & 2 & 1,606.20 \\
PI & 11 & 1 & 2,383.00 \\
RO & 14 & 2 & 4,762.20 \\
MS & 50 & 5 & 8,551.69 \\
RN & 56 & 5 & 9,992.60 \\
PB & 37 & 6 & 16,395.50 \\
MT & 144 & 4 & 17,007.82 \\
CE & 90 & 12 & 19,517.84 \\
\bottomrule
\end{tabular}
\caption{Lowest-Performing Sales Regions by Revenue}
\end{table}

The regions \texttt{AM}, \texttt{PA}, \texttt{SE}, \texttt{PI}, and
\texttt{RO} had the lowest total revenue. These regions may benefit from
additional marketing campaigns because their sales activity is much lower than
the strongest regions.

However, the results should be interpreted carefully. Some of these regions have
only one or two sellers, which suggests that low revenue may be caused by weak
market presence rather than only weak customer demand. Therefore, management
should consider both promotional campaigns and seller expansion strategies in
these regions.

For comparison, the strongest region was \texttt{SP}, with 78,604 units sold and
8,509,511.46 in total revenue. This shows a large performance gap between
\texttt{SP} and the lower-performing regions.


\subsection{Pricing and Discount Timing Analysis}

The third management question asks when products should be offered at full price
and when discounts should be used. The dataset does not contain a direct discount
field, so monthly sales volume and total revenue were used to identify stronger
and weaker sales periods.

High-demand months are better candidates for full-price selling because customers
are already purchasing more products. Low-demand months are better candidates for
discounts, promotions, or additional marketing because sales activity is weaker.

\begin{table}[H]
\centering
\begin{tabular}{lrr}
\toprule
Month & Units Sold & Total Revenue \\
\midrule
May & 11,814 & 1,466,882.94 \\
August & 11,939 & 1,393,276.34 \\
July & 11,379 & 1,349,557.98 \\
April & 10,396 & 1,314,203.77 \\
March & 10,914 & 1,312,555.10 \\
\bottomrule
\end{tabular}
\caption{Strongest Months by Total Revenue}
\end{table}

Based on this result, managers should consider selling products closer to full
price during March, April, May, July, and August. These months had the strongest
sales performance and may not require aggressive discounting.

\begin{table}[H]
\centering
\begin{tabular}{lrr}
\toprule
Month & Units Sold & Total Revenue \\
\midrule
September & 4,740 & 607,534.64 \\
October & 5,527 & 688,572.76 \\
December & 6,188 & 726,044.09 \\
November & 8,475 & 987,765.37 \\
January & 8,950 & 1,036,443.36 \\
\bottomrule
\end{tabular}
\caption{Weakest Months by Total Revenue}
\end{table}

Discounts and promotional campaigns should be considered during September,
October, December, November, and January because these months had weaker sales
performance. Promotions during these periods may help increase demand and improve
revenue.

The results should be interpreted carefully because the dataset does not cover
all calendar months equally. The data begins in September 2016 and ends in August
2018, so some periods may be incomplete. Even with this limitation, the monthly
trend provides useful guidance for pricing and promotional planning.


\section{Job Search Project 1-10}

\subsection{Research Method}

For this part of the project, I reviewed current job-posting results for business
positions in accounting, finance, sales, marketing, and human resources. The goal
was to find jobs that are not purely technical IT jobs, but still require
information systems knowledge. The selected jobs show that information systems
skills are now important across many business functions.

The research focused on three representative roles:

\begin{itemize}
    \item Accounting or finance analyst roles
    \item Sales operations analyst roles
    \item Human resources information systems analyst roles
\end{itemize}


\subsection{Job 1: Accounting or Finance Analyst}

One job category that requires information systems knowledge is accounting and
financial analysis. Current job postings for accounting and finance analyst roles
commonly require experience with ERP systems, accounting systems, advanced Excel,
large datasets, and business reporting tools.

For example, an Accounting Transaction Analyst posting listed experience with ERP
or accounting systems, comfort working with large datasets, and advanced Excel
skills as important qualifications. Other financial analyst postings also mention
SQL, Tableau, budgeting systems, and financial automation.

The information systems knowledge required for this type of job includes:

\begin{itemize}
    \item Understanding ERP and accounting information systems
    \item Using spreadsheet tools such as Excel for financial analysis
    \item Working with large datasets
    \item Creating reports for budgeting, forecasting, and performance analysis
    \item Understanding data accuracy and internal controls
\end{itemize}

This role connects directly to management information systems because financial
decisions depend on accurate, timely, and well-organized data. Instead of only
recording financial transactions, accounting and finance employees are expected
to use systems to analyze data and support decision making.


\subsection{Job 2: Sales Operations Analyst}

A second job category requiring information systems knowledge is sales operations.
A Sales Operations Analyst posting required familiarity with SAP, CRM systems,
SQL, Salesforce, Power BI, spreadsheet analysis, and sales reporting.

The information systems knowledge required for this type of job includes:

\begin{itemize}
    \item Understanding customer relationship management systems such as Salesforce
    \item Using ERP or SAP data for sales and operations reporting
    \item Writing or understanding SQL queries
    \item Creating dashboards in Power BI or similar business intelligence tools
    \item Analyzing sales performance, territory performance, and customer data
\end{itemize}

This role shows how information systems support sales strategy. Sales managers
need accurate data about customers, products, territories, and revenue. A sales
operations analyst uses information systems to turn raw sales data into reports
that help managers improve productivity, identify weak regions, and make better
marketing or sales decisions.


\subsection{Job 3: HRIS or Workday Analyst}

A third job category requiring information systems knowledge is human resources
information systems analysis. HRIS Analyst and Workday Analyst postings require
knowledge of Workday HCM, HRIS configuration, system administration, reporting,
dashboards, data accuracy, security, and process improvement.

The information systems knowledge required for this type of job includes:

\begin{itemize}
    \item Understanding HRIS, HRMS, or HCM systems
    \item Maintaining employee records in a human resources system
    \item Configuring Workday or similar HR platforms
    \item Creating HR reports and dashboards
    \item Protecting employee data and following privacy requirements
    \item Supporting payroll, benefits, recruiting, and employee lifecycle data
\end{itemize}

This role is important because human resources departments manage large amounts
of sensitive employee data. Information systems help HR employees maintain
accurate records, support payroll and benefits, track workforce changes, and
provide reports to managers.


\subsection{Preparation for These Jobs}

To prepare for these jobs, a student should develop both business knowledge and
information systems skills. The most important preparation steps are:

\begin{itemize}
    \item Learn spreadsheet analysis, including formulas, pivot tables, lookup
    functions, and charts.
    
    \item Learn database concepts, including tables, primary keys, foreign keys,
    joins, grouping, and SQL queries.
    
    \item Practice using business intelligence tools such as Power BI or Tableau.
    
    \item Understand major enterprise systems, including ERP, CRM, and HRIS
    platforms.
    
    \item Build small portfolio projects that show the ability to clean data,
    query data, analyze trends, and present results.
    
    \item Develop communication skills so technical findings can be explained to
    business managers.
\end{itemize}

This MIS project helps prepare for these jobs because it uses a real transactional
database, converts data into usable formats, applies SQL queries, and summarizes
results for management decision making.


\subsection{Summary of Job Search Findings}

The job search shows that information systems knowledge is required in many
business careers, not only in IT departments. Accounting and finance roles use
ERP systems, spreadsheets, and reporting tools. Sales operations roles use CRM,
SAP, SQL, and business intelligence dashboards. Human resources roles use HRIS
and Workday systems to manage employee information and reporting.

Overall, the most useful information systems skills for business students are
spreadsheet analysis, database querying, business intelligence reporting,
enterprise system knowledge, and data-driven decision making.


\subsection{Sources Consulted}

\begin{itemize}
    \item MUFG Accounting Transaction Analyst job posting:
    \url{https://mufgub.wd3.myworkdayjobs.com/en-US/MUFG-Careers/job/Accounting-Transaction-Analyst_10072732-WD}
    
    \item Monster financial analyst job search results:
    \url{https://www.monster.com/jobs/q-financial-analyst-jobs-l-san-francisco-ca}
    
    \item ZEISS Sales Operations Analyst job posting:
    \url{https://zeissgroup.wd3.myworkdayjobs.com/en-US/External/job/Analyst-II--Sales-Operations_JR_1049741}
    
    \item Monster HRIS Analyst job search results:
    \url{https://www.monster.com/jobs/q-hris-analyst-jobs}
    
    \item O*NET Human Resources Specialists:
    \url{https://www.onetonline.org/link/summary/13-1071.00}
    
    \item O*NET Market Research Analysts Technology Skills:
    \url{https://www.onetonline.org/link/tt/13-1161.00/43232306}
    
    \item BLS Market Research Analysts Occupational Outlook:
    \url{https://www.bls.gov/ooh/business-and-financial/market-research-analysts.htm}
\end{itemize}


\section{Conclusion}

This project shows how information systems can improve business performance,
management decision making, and operational control. The Snyder's of Hanover case
demonstrated the problems created by manual spreadsheet-based reporting. Slow
data collection, repeated manual entry, and inconsistent spreadsheet updates can
delay financial reporting and reduce the quality of management decisions. A
centralized financial information system or ERP system would help the company
produce faster, more accurate, and more useful reports.

The Dollar General case showed that inventory management problems should not be
solved by purchasing technology without first understanding the business need.
Before investing in an information system, management must define the problem,
identify required system features, evaluate costs and benefits, plan employee
training, and consider security requirements. An automated inventory management
system using barcode scanning and a centralized database could reduce inventory
losses and improve store-level decision making.

The database analysis project demonstrated how raw transactional data can be
converted into useful management information. SQL queries were used to analyze
product demand, regional sales performance, seller performance, and monthly sales
trends. The analysis found that product categories such as
\texttt{cama\_mesa\_banho}, \texttt{beleza\_saude}, \texttt{esporte\_lazer},
\texttt{moveis\_decoracao}, and \texttt{informatica\_acessorios} should be
prioritized for restocking because they had the highest sales volume. The
analysis also showed that lower-performing regions such as \texttt{AM},
\texttt{PA}, \texttt{SE}, \texttt{PI}, and \texttt{RO} may benefit from
additional marketing or seller expansion. Finally, the monthly sales analysis
showed that stronger sales months such as March, April, May, July, and August are
better candidates for full-price selling, while weaker months such as September,
October, December, November, and January may require discounts or promotional
campaigns.

The job search project showed that information systems knowledge is important in
many business careers, not only in technical IT positions. Accounting, finance,
sales, marketing, and human resources jobs increasingly require skills in
spreadsheets, databases, ERP systems, CRM systems, HRIS platforms, SQL, and
business intelligence tools.

Overall, this project demonstrates that information systems help organizations
collect data, organize data, analyze trends, and make better decisions. Managers
who understand information systems can improve efficiency, reduce errors, control
costs, identify opportunities, and respond more quickly to business problems.


\newpage
\appendix

\section{Appendix A: SQLite to CSV Export Script}

The following Python script was used to export the SQLite database tables into
CSV files. This made the data easier to inspect and made it possible to use the
data in spreadsheet or database software if needed.

\lstinputlisting[
    style=codeStyle,
    language=Python,
    caption={SQLite to CSV Export Script}
]{../scripts/sqlite_to_csv.py}


\newpage
\section{Appendix B: Database Profile and Base View SQL}

The following SQL file was used to inspect the database, count records, check the
date range, and create the main joined sales analysis view.

\lstinputlisting[
    style=codeStyle,
    language=SQL,
    caption={Database Profile and Sales Analysis Base View}
]{../sql/sales_analysis_queries.sql}


\newpage
\section{Appendix C: Restocking Analysis SQL}

The following SQL file was used to identify high-demand product categories and
products that should be prioritized for restocking.

\lstinputlisting[
    style=codeStyle,
    language=SQL,
    caption={Restocking Analysis SQL Queries}
]{../sql/restock_analysis.sql}


\newpage
\section{Appendix D: Promotional Campaign Analysis SQL}

The following SQL file was used to identify low-performing regions and sellers
that may benefit from promotional campaigns or additional marketing support.

\lstinputlisting[
    style=codeStyle,
    language=SQL,
    caption={Promotional Campaign Analysis SQL Queries}
]{../sql/promotion_analysis.sql}


\newpage
\section{Appendix E: Pricing and Discount Timing SQL}

The following SQL file was used to analyze monthly sales trends and identify
stronger and weaker sales periods for pricing and promotion decisions.

\lstinputlisting[
    style=codeStyle,
    language=SQL,
    caption={Pricing and Discount Timing SQL Queries}
]{../sql/pricing_timing_analysis.sql}


\end{document}